Study of Intra- and Inter-Job Interference with Different Job Allocations

### 2015-10-16 submission.tex
update some todos, fixup some cites

Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability. Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance. In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required. In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference. By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios. \changed{Our comprehensive analysis in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.} \TODO{We propose a flexible allocation strategy? If not, we want to weaken this, and instead say something like the results of this research illuminate a path towards access-pattern-aware HPC job allocation.}

@@

The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research. Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity. Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite doesn't fit here, should put somewhere else}. To be effective, they must maintain high system utilization. These communication and I/O intensive jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources. As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.

@@

On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used. The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}. The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{wrong cite - this implies that Zhou designed the algorithm or the BG :) }. The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies. However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible. On the other side of the coin, the non-contiguous allocation strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality. Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation. While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.} \TODO{a cite would be good here.}

@@

In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500} \TODO{need to say which date this applies to}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although our analyses are based on torus network, the ideas conveyed in this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{is there a better cite for this?}. We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)  \cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment. We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations. We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.

\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
The observations of this work are the following:
\begin{comment}

@@

There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns. In this work, we select three representative applications from the DOE Design Forward Project \cite{designforwardwebpage} \TODO{better cite for this?}. Each application exhibits a distinctive communication pattern that is commonly seen HPC applications. We believe these three applications can represent a big group of applications that conform to similar communication patterns. The applications studied in this work are the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter. The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data from publicly available traces~\cite{designforwardwebpage}.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
        \caption{Communication Topology}
        \label{fig:amg-communication-topology}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
        \caption{Data Volume}
        \label{fig:amg-data-trans}
    \end{subfigure}
    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in AMG.}
\end{figure}

@@

Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks. Note that the dominant communication pattern of the application doesn't change with scale. We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks. Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}} \TODO{need better cites for these...}.

@@

The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}} \TODO{should cite the nek5000 website. couldn't find a canonical citation for the software}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

@@

MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}} \TODO{cite?}. MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle. It is widely used for structured grid physics packages.

@@

\bibliographystyle{abbrv}
\bibliography{./reference}

>>>>

Network contention between concurrently running jobs on HPC systems is a primary cause of performance variability. Optimizing job allocation and avoiding network sharing are hence crucial to alleviate the potential degradation in performance. In order to do these effectively, an understanding of the interaction between applications, their communication patterns, and the system architecture are required. In this work, we choose three representative HPC applications from DOE Design Forward Project and conduct detailed simulations of high-performance network topologies to analyze both intra- and inter-job interference. By scrutinizing the communication behaviors of these applications, we identify relationships between these behaviors and the possible interference induced by different job allocation scenarios. \changed{Our comprehensive analyses in this work illuminate a path towards job-communication-pattern-aware HPC resource allocation.}

@@

The scale of supercomputers continually grows at a breakneck pace to accommodate the computing power requirements for scientific research. Supercomputers have \changed{tens of thousands} of nodes  and serve as an irreplaceable research vehicle for scientific problems with increasing size and complexity. Supercomputers are usually employed as a shared resource to accommodate many parallel applications (jobs) running concurrently \changed{\cite{jose-ipdps15}} \TODO{cite (still) doesn't fit here. I don't think you need to defend this sentence with a cite}. To be effective, they must maintain high system utilization. These communication and I/O intensive jobs share the system infrastructure such as network and I/O bandwidth, and inevitably there is contention over these shared resources. As supercomputers continue to evolve, these shared resources are increasingly the bottleneck for performance.

@@

On the widely used torus-connected HPC systems \cite{bgq,tofu,titan}, two allocation strategies are commonly used. The contiguous allocation strategy assigns to each job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}. The partition-based allocation approach used in Blue Gene series systems is an example of such a strategy \changed{\cite{bgloverview}} \TODO{this one is OK. Tried seeing if there were a more recent cite for this, but didn't come up with anything}. The contiguous strategy favors application performance through isolated networking within a partition and the locality that implies. However, this strategy can cause both internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization than is otherwise possible. On the other side of the coin, the non-contiguous allocation strategy, used by the Cray XT/XE series \cite{carl-cug}, assigns free nodes to jobs regardless of contiguity, though of course efforts are made to maximize locality. Figure~\ref{fig:overview_sub2} shows an example non-contiguous allocation. While eliminating internal and external fragmentation as seen in contiguous allocation systems, in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones \changed{\cite{abhinav-sc13}.}

@@

In this work, we focus on an in-depth analysis of intra- and inter-job communication interference with different job allocation strategies on torus-connected HPC systems, which are used on six of the top 10 supercomputers on the \changed{June 2015} Top500 lists~\cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, such as Mira at Argonne Nation Laboratory and Sequoia at Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the ``Tofu'' system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although our analyses are based on torus networks, the ideas conveyed in this work is applicable to networks with different topologies.

We selected three signature applications from the DOE Design Forward Project~\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns \TODO{checked with some folks and there actually isn't a better citation for now.}. We use a sophisticated simulation toolkit named CODES (standing for Co-Design of Multi-layer Exascale Storage Architectures)~\cite{Jason-2011} as a research vehicle to evaluate the performance of these applications with various allocations in a controlled environment. We then analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations. We believe the insights presented in this work can be very useful for the design of future HPC batch job schedulers and resource managers.

\NOTE{JJ removed the list of contributions, because they are exactly the same as the preceding paragraph.}
\begin{comment}
The observations of this work are the following:

@@

There are many applications running on HPC systems, such as Mira \cite{bgq} and Titan \cite{titan}, that are communication intensive and conform to various communication patterns. In this work, we select three representative applications from the DOE Design Forward Project. Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. We believe that the communication patterns of these applications are representative of a wide array of applications run on leadership-class machines. Specifically, we study the Algebraic MultiGrid Solver (AMG), Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. The communication pattern figures we present in Section~\ref{sec:amg}, \ref{sec:crystalrouter}, and \ref{sec:multigrid} are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.24\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_pip}
        \caption{Communication Topology}
        \label{fig:amg-communication-topology}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.24\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/appstudy/amg/amg_data_transfer}
        \caption{Data Volume}
        \label{fig:amg-data-trans}
    \end{subfigure}
    \caption{AMG. (a) rank-to-rank communication topology graph. (b) the data sent/received by each rank in AMG.}
\end{figure}

@@

Figure~\ref{fig:amg-communication-topology} shows the communication topology of a small scale AMG execution with 216 MPI Ranks. Note that the dominant communication pattern of the application doesn't change with scale. We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, each rank has intensive communication with up to six neighbors, depending on the boundaries of the ranks. Applications with similar pattern include PARTISN \changed{\cite{partisn}} and SNAP \changed{\cite{snap}}.

@@

The second MiniApp covered is CrystalRouter, the extracted communication kernel of the full application Nek5000~\cite{crystalrouter}, which is a spectral element CFD application developed at Argonne National Laboratory \changed{\cite{nek5000}}. It features spectral element multi-grid solvers coupled with a highly scalable, parallel coarse-grid solver that is widely used for projects including ocean current modeling, thermal hydraulics of reactor cores, and spatiotemporal chaos. CrystalRouter demonstrates the ``many-to-many'' communication pattern through a scalable multi-stage communication process. The way CrystalRouter works is nodes compute for a while, then synchronize and communicate, continually alternating between these two types of activities.

@@

MultiGrid is geometric multi-grid v-cycle from the production elliptic solver BoxLib, a software framework for massively parallel block-structured adaptive mesh refinement (AMR) codes \changed{\cite{boxlib}}. MultiGrid conforms to many-to-many communication pattern with decreasing message size and collectives for different parts of the multi-grid v-cycle. It is widely used for structured grid physics packages.

@@

\bibliographystyle{IEEEtran}
\bibliography{./reference}
